\begin{thebibliography}{Abramovitz\thinspace$\&$\thinspace Stegun}

\bibitem[Abramovitz\thinspace$\&$\thinspace Stegun]{AbrSteg}
Milton Abramovitz and Irene A. Stegun, {\em Handbook of Mathematical Functions}, Dover publications, Inc., 1964.

\bibitem[Achtley\thinspace$\&$\thinspace Bryant]{AtchBry}
William R. Achtley and Edwin H. Bryant editors,
{\em Benchmark Papers in Systematic and Evolutionary Biology},
Vol. 1, Dowden, Hutchinson\thinspace$\&$\thinspace Ross, Inc.,
Stroudsburg, Pa.; distributed by Halsted Press [John
Wiley\thinspace$\&$\thinspace Sons, Inc.], New York, 1975.

\bibitem[Bass]{Bass}J. Bass, {\em Cours de Mathématiques}, Tome II, Masson, 1968.

\bibitem[Beck]{Beck}Kent Beck, {\em Smalltalk Best Practice Patterns}, Prentice
Hall, 1997.

\bibitem[Berry\thinspace$\&$\thinspace Linoff]{BerLin}
Michael J.A. Berry and Gordon Linoff, {\em Data mining for
marketing, sales and customer support}, John
Wiley\thinspace$\&$\thinspace Sons, Inc., 1997.

\bibitem[Cormen {\it et al.}]{CorLeiRiv}
Thomas H. Cormen, Charles E. Leiserson and Ronald L. Rivest, {\em
Introduction to Algorithms}, McGraw-Hill, 1990.

\bibitem[Gamma {\it et al.}]{GoF}
Erich Gamma, Richard Helm, Ralph Johnson and John Vlissides , {\em
Design Patterns}, Addison-Wesley, 1995.

\bibitem[Gullberg]{Gullberg} Jan Gullberg, {\em Mathematics From the Birth of the Numbers},
W.W. Norton\thinspace$\&$\thinspace Company, 1997.

\bibitem[Ifrah]{Ifrah}Georges Ifrah, {\em Histoire Universelle des Chiffres},
Robert Laffont, 1994.

\bibitem[Knudth 1]{Knuth1}Donald E. Knuth, {\em The Art of
Computer Programming} Vol. 1, Addison-Wesley, 1973.

\bibitem[Knudth 2]{Knuth2}Donald E. Knuth, {\em The Art of Computer Programming} Vol. 2,
Addison-Wesley, 1981.

\bibitem[Knudth 3]{Knuth3}Donald E. Knuth, {\em The Art of
Computer Programming} Vol. 3, Addison-Wesley, 1973.

\bibitem[Koza {\it et al.}]{Koza}John R. Koza, Forrest H, Bennett III, David Andre and Martin A. Keane,
{\em Genetic Programming III}, Morgan Kaufmann, 1999.

\bibitem[Law\thinspace$\&$\thinspace Kelton]{LawKel}
Averill M. Law and W. David Kelton, {\em Simulation Modeling and
Analysis}, McGraw-Hill, 1982.

%\bibitem[Lin\thinspace$\&$\thinspace Lee]{LinLee}
%Chin-Teng Lin and C.S.
%George Lee, {\em Neural Fuzzy Systems}, Prentice Hall, 1996.

\bibitem[Phillips\thinspace$\&$\thinspace Taylor]{PhiTay}
G.M. Phillips and P.J. Taylor, {\em Theory and Applications of
Numerical Analysis}, Academic Press: London and New York, 1973.

\bibitem[Press {\it et al.}]{Press}William H. Press, Saul A. Teukolsky, William T. Vetterling
and Brian P. Flannery, {\em Numerical recipes for C : the art of
scientific computing}, Cambridge University Press, 1992.

\bibitem[Alpert {\it et al.}]{StDesPat}Sherman R. Alpert, Kyle Brown and Bobby Woolf,
{\em Design Pattern Smalltalk Companion}, Addison-Wesley, 1998.

\bibitem[Smith]{Smalltalk} David N. Smith,
{\em IBM Smalltalk, The language}, Addison-Wesley, 1995.

\bibitem[Flanagan]{Java} David Flanagan,
{\em Java in a nutshell}, O'Reilly, 1996.

\end{thebibliography}
